Signatures and Induction Principles for Higher Inductive-Inductive Types

### 2019-12-02 paper.tex

Note that we can assume arbitrary structures in the external $\hat{\Gamma}$
context, which in particular allows us to specify HIITs depending on other
(external) HIITs. We can do this by first specifying a HIIT, then using the
methods in this paper to compute notions of algebras and induction, then assume
the HIIT in the external context of another HIIT signature.

@@

(7) $\U$ is also closed under a function space where the domain is an
external type and the codomain is a small source theory type. We
overload the application notation for external parameters, as it is
usually clear from context which application is meant. The rules allow
types with infinitary constructors, for example, trees branching by a
possibly infinite external type $\hat{A}$:
\[
\hat{A}\in\Type_{\hat{0}}\vdash\,\boldsymbol{\cdot},\,\,\,T:\U,\,\,\,leaf:\underline{T},\,\,\,node:((\hat{x}\in \hat{A})\ra T)\ra\underline{T}
\]
Here, $node$ has a function type (3) with a function type (7) in the
domain. More generally, we can define $W$-types
\cite{abbot05containers} as follows. $\hat{S}$ describes the
``shapes'' of the constructors and $\hat{P}$ the ``positions'' where
recursive arguments can appear.
\[
\hat{S}\in\Type_{\hat{0}}\commah\hat{P}:\hat{S}\rah\Type_{\hat{0}}\vdash\,\boldsymbol{\cdot},\,\,\,W:\U,\,\,\,sup: (\hat{s} \in \hat{S})\ra((\hat{p} \in \hat{P}\,\hat{s})\ra W)\ra \underline{W}
\]
For a more complex infinitary example, see the definition of Cauchy
reals in \cite[Definition 11.3.2]{HoTTbook}. It can be also found as
an example file in our Haskell implementation.

@@

The specification of the $\blank^\M$ interpretation is as following.
\[
\infer{\hat{\Gamma}\vdashh\Delta^\M \in \Delta^\A\rah\Delta^\A\rah \Type_{\hat{i}}}
      {\hat{\Gamma}\vdash\Delta}
\hspace{1em}
\infer{
  \hat{\Gamma}\vdashh A^\D \in \hat{(}\hat{\gamma_0\,\gamma_1}\in\Delta^\A\hat{)}\rah \Delta^\M\,\hat{\gamma_0\,\gamma_1}\rah  A^\A\,\hat{\gamma_0}\rah A^\A\,\hat{\gamma_1}\rah \Type_{\hat{i}}}
      {\hat{\Gamma}\semicol\Delta\vdash A}
\]

>>>>

Note that we can assume arbitrary structures in the external $\hat{\Gamma}$
context, which in particular allows us to specify HIITs depending on other
(external) HIITs. We can do this by first specifying a HIIT, then using the
translations in Sections \ref{sec:algebras}-\ref{sec:sections} to compute
notions of algebras and induction, then assume the HIIT in the external context
of another HIIT signature.

@@

(7) $\U$ is also closed under a function space where the domain is an
external type and the codomain is a small source theory type. We
overload the application notation for external parameters, as it is
usually clear from context which application is meant. The rules allow
types with infinitary constructors, for example, trees branching by a
possibly infinite external type $\hat{A}$:
\[
\hat{A}\in\Type_{\hat{0}}\vdash\,\boldsymbol{\cdot},\,\,\,T:\U,\,\,\,leaf:\underline{T},\,\,\,node:((\hat{x}\in \hat{A})\ra T)\ra\underline{T}
\]
Here, $node$ has a function type (3) with a function type (7) in the
domain. More generally, we can define $W$-types
\cite{abbot05containers} as follows. $\hat{S}$ describes the
``shapes'' of the constructors and $\hat{P}$ the ``positions'' where
recursive arguments can appear.
\[
\hat{S : \Type_0, P : S \ra \Type_0} \vdash\,\boldsymbol{\cdot},\,\,\,W:\U,\,\,\,sup: (\hat{s} \in \hat{S})\ra((\hat{p} \in \hat{P}\,\hat{s})\ra W)\ra \underline{W}
\]
For a more complex infinitary example, see the definition of Cauchy
reals in \cite[Definition 11.3.2]{HoTTbook}. It can be also found as
an example file in our Haskell implementation.

@@

The specification of the $\blank^\M$ interpretation is as following.
\[
\infer{\hat{\Gamma}\vdashh\Delta^\M \in \Delta^\A\rah\Delta^\A\rah \Type_{\hat{i}}}
      {\hat{\Gamma}\vdash\Delta}
\hspace{1em}
\infer{
  \hat{\Gamma}\vdashh A^\M \in \hat{(}\hat{\gamma_0\,\gamma_1}\in\Delta^\A\hat{)}\rah \Delta^\M\,\hat{\gamma_0\,\gamma_1}\rah  A^\A\,\hat{\gamma_0}\rah A^\A\,\hat{\gamma_1}\rah \Type_{\hat{i}}}
      {\hat{\Gamma}\semicol\Delta\vdash A}
\]
